Dynamical modeling of tidal streams

### 2013-12-31 tex/ms.tex
streams: typo
streams: first section of the discussion

Similar procedures can be followed to evaluate other projections and
marginalizations of the PDF, \eg, $p(b|l)$ for the observed track of
the stream on the sky, to convolve the PDF with observational
uncertainties, and to calculate moments of the PDF, \eg, the density
and velocity dispersion along the stream. Another example is shown in
\figurename~\ref{fig:gd1_pdf_l}. This Figure shows the model's density
as a function of Galactic longitude and compares it to that of the
simulated stream in the final snapshot (\ie, that of
\figurename~s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv}). The model
density here is calculated using a Monte Carlo mock stream sample,
drawn as described in \sectionname~\ref{sec:mock}, rather than through
direct numerical integration.

@@

@@

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
  Binney,~J. \& Tremaine,~S. 2008, Galactic Dynamics: Second Edition
\bibitem[Binney(2008)]{Binney08a}
  Binney,~J. 2008, \mnras, 386, L47
\bibitem[Binney(2012)]{Binney12a}
  Binney,~J. 2012, \mnras, 426, 1324
\bibitem[Bovy \etal(2011)]{Bovy11a}
  Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T. 2011, Ann. Appl. Stat. 5, 1657, arXiv:0905.2979v1
\bibitem[Carlberg(2012)]{Carlberg12a}
Carlberg,~R.~G. 2012, \apj, 748, 20
\bibitem[Dehnen(2000)]{Dehnen00a}
  Dehnen,~W. 2000, \apj, 536, L39
\bibitem[Dehnen(2002)]{Dehnen02a}
  Dehnen,~W. 2002, J.~Comput.~Phys., 179, 27
\bibitem[Eyre \& Binney(2011)]{Eyre11a}
  Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
\bibitem[Gilks \& Wild(1992)]{Gilks92a}
  Gilks,~W.~R. \& Wild,~P. 1992, Applied Statistics, 41, 337
\bibitem[Helmi \& White(1999)]{Helmi99a}
  Helmi,~A. \& White,~S.~D.~M. 1999, \mnras, 307, 495
\bibitem[Johnston \etal(1999)]{Johnston99a}
  Johnston,~K.~V., Zhao,~H., Spergel,~D.~N., \& Hernquist,~L. 1999, \apj, 512, L109
\bibitem[Koposov \etal(2010)]{Koposov10a}
  Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
\bibitem[McGill \& Binney(1990)]{McGill90a}
  McGill,~C. \& Binney,~J. 1990, \mnras, 244, 634
\bibitem[McMillan \& Binney(2008)]{McMillan08a}
  McMillan,~P.~J. \& Binney,~J.~J. 2008, \mnras, 390, 429
\bibitem[Ngan \& Carlberg(2014)]{Ngan14a}
  Ngan,~W.~H.~W. \& Carlberg,~R.~G. 2014, \apj, submitted
\bibitem[Pe{\~n}arrubia \etal(2012)]{Penarrubia12a}
  Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
\bibitem[Price-Whelan \& Johnston(2013)]{PriceWhelan13a}
  Price-Whelan,~A.~M. \& Johnston,~K.~V.~2013, \apj, 778, L12
\bibitem[Sanders(2012)]{Sanders12a}
  Sanders,~J. 2012, \mnras, 426, 128
\bibitem[Sanders \& Binney(2013a)]{Sanders13a}
  Sanders,~J. \& Binney,~J. 2013a, \mnras, 433, 1813
\bibitem[Sanders \& Binney(2013b)]{Sanders13b}
  Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
\bibitem[Shoemake(1985)]{Shoemake85a}
  Shoemake,~K. 1985, ACM SIGGRAPH Computer Graphics 3, 45
\bibitem[Teuben(1995)]{Teuben95a}
  Teuben,~P.~J. 1995, in Astronomical Data Analysis Software and Systems IV, 
  ed.~R.~Shaw, H.~E.~Payne and J.~J.~E.~Hayes., PASP Conf Series 77, 398
\bibitem[Tremaine(1999)]{Tremaine99a}
  Tremaine,~S. 1999, \mnras, 307, 877
\bibitem[Yoon \etal(2011)]{Yoon11a}
  Yoon,~J.~H., Johnston,~K.~V., \& Hogg,~D.~W. 2011, \apj, 731, 58
\end{thebibliography}

>>>>

Similar procedures can be followed to evaluate other projections and
marginalizations of the PDF, \eg, $p(b|l)$ for the observed track of
the stream on the sky, to convolve the PDF with observational
uncertainties, and to calculate moments of the PDF, \eg, the density
and velocity dispersion along the stream. Another example is shown in
\figurename~\ref{fig:gd1_pdf_l}. This Figure shows the model's density
as a function of Galactic longitude and compares it to that of the
simulated stream in the final snapshot (\ie, that of
\figurename s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv}). The model
density here is calculated using a Monte Carlo mock stream sample,
drawn as described in \sectionname~\ref{sec:mock}, rather than through
direct numerical integration.

\subsection{Dynamical fitting}

There are multiple existing methods for fitting data on tidal streams
with models for the Milky Way potential, however, none of these allow
for the likelihood of a model to be evaluated using a well-defined
generative model. The only exceptions to this are orbit-fitting
methods, but these use a faulty generative model (see below), and the
method of \citet{Varghese11a}. The latter is in many ways similar to
the method proposed here, in that it calculates the stream track and
uses this as the basis of the stream inference. They calculate the
track directly in position--velocity space by orbit integration of
stream members released at the Lagrange points. However, they do not
relate the observed width of the stream to the velocity dispersion of
the progenitor---and thus the location of the Lagrange points and the
position of the track---and it is therefore only weakly constrained
and has to be assumed. In general, it is straightforward to write down
generative models of tidal streams in position--velocity space by
substituting our initial offsets $(\Delta \veco,\Delta \veca)$ by
similar offsets $(\Delta \vecx,\Delta \vecv)$ and integrating both the
progenitor and offset stream members forward in time. However,
evaluating such models when fitting observational data requires large
numbers of orbit integrations to marginalize over stripping time
(which cannot be performed analytically in this case) and missing
phase--space dimensions. Thus, any such method will likely be orders
of magnitude slower than the method put forth here.

The forward model of this paper is superior to all other
stream-fitting methods in that it allows easy marginalization over
missing or noisily measured phase--space coordinates. The obvious
exception to this is orbit fitting, which as a purely one-dimensional
model is easy to integrate over. However, orbit fitting likely
produces biased results as streams do not delineate single orbits over
their full angle widths (see below). The fiducial model I propose in
this paper allows for a straightforward replacement of orbit fitting,
at the expense of adding two extra parameters, $\sigv$ and $t_d$, in
addition to the six-dimensional position of the progenitor. The orbit
in orbit fitting then gets replaced by the stream track, calculated as
in \sectionname~\ref{sec:track}. To constrain $\sigv$ and $t_d$ it is
necessary to also use the observed width and length of the
stream. While cold streams are typically too narrow to be resolved in
$D$ or any of the velocity components, the observed width of the
stream's sky projection can be measured and used as a
constraint. Future high-resolution spectroscopic data,
$\mu\mathrm{as}$-level astrometric data, or percent-level distances
for standard candles may allow the width and structure of tidal
streams to be resolved and provide further constraints on the
generative model. The switch from orbit fitting to stream-track
fitting comes at the expense of $\approx100$ orbit integrations for
each model rather than a single integration. If the progenitor is
unknown we also need to marginalize over whether we are seeing a
leading or a trailing stream.

The dynamical framework proposed here can be used in situations where
fitting a stream track is insufficient, for exampe, when high-quality
data in some dimensions on individual stars are available. Many of the
approaches in the literature do not require a model progenitor
position. This includes methods that attempt to minimize the spread of
orbital energies or actions in the stream (\eg,
\citealt{Binney08a,Penarrubia12a}, Sanderson \etal, 2014, in
preparation) as well as the method of \citet{Sanders13b}, which
constrains the potential by requiring angle differences in the stream
to lie along the same direction as frequency differences (see
\equationname~\ref{eq:pdf} and subsequent discussion). These methods
require good measurements of the phase--space coordinates of stream
members, or at the very least, \emph{some} measurement of all of the
six phase--space coordinates (although the latter can be relaxed when
excellent measurements of the line-of-sight velocity along the stream
are available; \citealt{Binney08a}). If such high-quality data are
present, these methods can provide good initial guesses for the PDF of
potential parameters that fit a stream, which can subsequently be used
in a more thorough exploration of the PDF and structure of the stream
using the framework of \sectionname~\ref{sec:pdf}. That these methods
work also goes to show that the progenitor parameters---phase--space
position, $\sigv$ and $t_d$---are not that important for the
constraints on the gravitational potential.

@@

\subsection{Do streams follow orbits?}

\subsection{Constraining progenitor properties with stream kinematics}

\subsection{Baseline models for stream-gap finding}

@@

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
  Binney,~J. \& Tremaine,~S. 2008, Galactic Dynamics: Second Edition
\bibitem[Binney(2008)]{Binney08a}
  Binney,~J. 2008, \mnras, 386, L47
\bibitem[Binney(2012)]{Binney12a}
  Binney,~J. 2012, \mnras, 426, 1324
\bibitem[Bovy \etal(2011)]{Bovy11a}
  Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T. 2011, Ann. Appl. Stat. 5, 1657, arXiv:0905.2979v1
\bibitem[Carlberg(2012)]{Carlberg12a}
Carlberg,~R.~G. 2012, \apj, 748, 20
\bibitem[Dehnen(2000)]{Dehnen00a}
  Dehnen,~W. 2000, \apj, 536, L39
\bibitem[Dehnen(2002)]{Dehnen02a}
  Dehnen,~W. 2002, J.~Comput.~Phys., 179, 27
\bibitem[Eyre \& Binney(2011)]{Eyre11a}
  Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
\bibitem[Gilks \& Wild(1992)]{Gilks92a}
  Gilks,~W.~R. \& Wild,~P. 1992, Applied Statistics, 41, 337
\bibitem[Helmi \& White(1999)]{Helmi99a}
  Helmi,~A. \& White,~S.~D.~M. 1999, \mnras, 307, 495
\bibitem[Johnston \etal(1999)]{Johnston99a}
  Johnston,~K.~V., Zhao,~H., Spergel,~D.~N., \& Hernquist,~L. 1999, \apj, 512, L109
\bibitem[Koposov \etal(2010)]{Koposov10a}
  Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
\bibitem[McGill \& Binney(1990)]{McGill90a}
  McGill,~C. \& Binney,~J. 1990, \mnras, 244, 634
\bibitem[McMillan \& Binney(2008)]{McMillan08a}
  McMillan,~P.~J. \& Binney,~J.~J. 2008, \mnras, 390, 429
\bibitem[Ngan \& Carlberg(2014)]{Ngan14a}
  Ngan,~W.~H.~W. \& Carlberg,~R.~G. 2014, \apj, submitted
\bibitem[Pe{\~n}arrubia \etal(2012)]{Penarrubia12a}
  Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
\bibitem[Price-Whelan \& Johnston(2013)]{PriceWhelan13a}
  Price-Whelan,~A.~M. \& Johnston,~K.~V.~2013, \apj, 778, L12
\bibitem[Sanders(2012)]{Sanders12a}
  Sanders,~J. 2012, \mnras, 426, 128
\bibitem[Sanders \& Binney(2013a)]{Sanders13a}
  Sanders,~J. \& Binney,~J. 2013a, \mnras, 433, 1813
\bibitem[Sanders \& Binney(2013b)]{Sanders13b}
  Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
\bibitem[Shoemake(1985)]{Shoemake85a}
  Shoemake,~K. 1985, ACM SIGGRAPH Computer Graphics 3, 45
\bibitem[Teuben(1995)]{Teuben95a}
  Teuben,~P.~J. 1995, in Astronomical Data Analysis Software and Systems IV, 
  ed.~R.~Shaw, H.~E.~Payne and J.~J.~E.~Hayes., PASP Conf Series 77, 398
\bibitem[Tremaine(1999)]{Tremaine99a}
  Tremaine,~S. 1999, \mnras, 307, 877\
\bibitem[Varghese \etal(2011)]{Varghese11a}
  Varghese,~A., Ibata,~R., \& Lewis,~G.~F. 2011, \mnras, 417, 198
\bibitem[Yoon \etal(2011)]{Yoon11a}
  Yoon,~J.~H., Johnston,~K.~V., \& Hogg,~D.~W. 2011, \apj, 731, 58
\end{thebibliography}
